---
title: "Price per Mile / Revenue Correlation"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
library(hexbin)
```

### 1 - Seek correlation between monthly lagged price per mile and number of Traveling Passengers

```{r ppm_cor, include=TRUE, echo=FALSE}
basepath<-"/Users/bertrand/snowflake/airfare_prediction"

dfa <- try(read.csv( paste0(basepath,"/data/B6_finance_ppm.csv") , header=TRUE, sep = "," ,stringsAsFactors=FALSE) )

dfa$Revenue.passengers <- as.numeric(as.character(dfa$Revenue.passengers))
dfa$Average.stage.length..miles. <- as.numeric(as.character(dfa$Average.stage.length..miles.))

pairs(dfa[,c(4,5,8,9,10,11,13,15,16,17,18,19,20)], pch = 19)

```

#### Conclusion:
Price purpase on 4, 6, 15, 21 and 30 days prior de departure have the best correlation with number of passengers.

Next, checking those variables for wide enough spread to be significant and avoid overfitting.

### 2 - Box charts price per mile variation per departure lag 
```{r withplotly, echo=FALSE}

plot_ly(dfa, x = ~Day_0, name='Day 0' ,type = "box") %>%
  add_trace(dfa, x = ~Day_1, name='Day 1' ,type = "box") %>%
  add_trace(dfa, x = ~Day_2, name='Day 2' ,type = "box") %>%
  add_trace(dfa, x = ~Day_3, name='Day 3' ,type = "box") %>%
  add_trace(dfa, x = ~Day_4, name='Day 4' ,type = "box") %>%
  add_trace(dfa, x = ~Day_5, name='Day 5' ,type = "box") %>%
  add_trace(dfa, x = ~Day_6, name='Day 6' ,type = "box") %>%
  add_trace(dfa, x = ~Day_7, name='Day 7' ,type = "box") %>%
  add_trace(dfa, x = ~Day_15, name='Day 15' ,type = "box") %>%
  add_trace(dfa, x = ~Day_21, name='Day 21' ,type = "box") %>%
  add_trace(dfa, x = ~Day_30, name='Day 30' ,type = "box") %>%
  add_trace(dfa, x = ~Day_60, name='Day 60' ,type = "box") %>%
  add_trace(dfa, x = ~Day_90, name='Day 90' ,type = "box") %>%
  add_trace(dfa, x = ~Day_180, name='Day 180' ,type = "box") %>%
  layout(title = "ppm ( price per mile ) in US Dollar")

```

#### Conclusion:
Keep Day 4, 15, 21 and 30 but prefer Day 6 over Day 7 for its variablility


### 3 - Building number of passengers predictive model based on ppm for prior booking on Day 4, 5, 7, 21 and 30
```{r monthly_predict, echo=FALSE}
fit <- lm(Revenue.passengers ~ Day_4+Day_6+Day_15+Day_21+Day_30 , data = dfa )
summary(fit)
pre <- predict( fit, dfa) 
dfa$pre <- pre
acu <- data.frame(cbind(actuals=dfa$Revenue.passengers, predicteds=pre))

#acu
plot(acu)
coef(fit)

bin<-hexbin(dfa$Revenue.passengers, pre, xbins=4) 
plot(bin, main="Hexagonal Binning")

```

#### Conclusion:
Model plotted above has a very significant predictive model. Next step predicting Quarterly Revenue.
Note: It is possible that this airline has a very predictive seasonal traffic that would explain such a blattant correlation.

## 4 - Predicting Quarterly Revenue from price per mile variation

```{r quarterly, echo=TRUE}
dfa <- try(read.csv( paste0(basepath,"/data/B6_quarterly_finance_ppm.csv") , header=TRUE, sep = "," ,stringsAsFactors=FALSE) )

dfa$Revenue <- as.numeric(as.character(dfa$Revenue))
dfa$Quarter_ending <- as.Date(dfa$Quarter_ending, "%m/%d/%y")
#pairs(dfa[,c(2,4,5,8,9,10,11,13,15,16,17)], pch = 19)

fit <- lm(Revenue ~ Day_4+Day_7+Day_15+Day_30+Day_21, data = dfa )
summary(fit)
pre <- predict( fit, dfa) 
dfa$pre <- pre
acu <- data.frame(cbind(actuals=dfa$Revenue, predicteds=pre))

acu
plot(acu)
coef(fit)

bin<-hexbin(dfa$Revenue, pre, xbins=3) 
plot(bin, main="Hexagonal Binning")

```

#### Conclusion: 
  >>> PRETTY GOOD, NO?

### 5 - Visual confirmation chart
Chart below is representing normalized (around their means) the fields: revenue, prediction, day 0 to 180 
```{r verif, echo=FALSE}
plot_ly(dfa , x = ~Quarter_ending, y=~Revenue/mean(dfa$Revenue)*100, type = "scatter", name = 'Revenue', mode='lines') %>%
  add_trace(y = ~pre/mean(dfa$pre)*100, name = 'Prediction', mode = 'lines+markers') %>%
  add_trace(y = ~Day_0/mean(dfa$Day_0)*100, name = 'Day_0', mode = 'lines+markers') %>%
  add_trace(y = ~Day_5/mean(dfa$Day_5)*100, name = 'Day_5', mode = 'lines+markers') %>%
  add_trace(y = ~Day_7/mean(dfa$Day_7)*100, name = 'Day_7', mode = 'lines+markers') %>%
  add_trace(y = ~Day_15/mean(dfa$Day_15)*100, name = 'Day_15', mode = 'lines+markers') %>%
  add_trace(y = ~Day_21/mean(dfa$Day_21)*100, name = 'Day_21', mode = 'lines+markers') %>%
  add_trace(y = ~Day_30/mean(dfa$Day_30)*100, name = 'Day_30', mode = 'lines+markers') %>%
  add_trace(y = ~Day_180/mean(dfa$Day_180)*100, name = 'Day_180', mode = 'lines+markers')


```


Note: Time serie is per quarter... January 2018 is in fact 2017 Q4 ...
